---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

by W L Man.  
31 July 2016.    

This is a Coursera Assignment for week 2 on Reproducible Research.  

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.    

The data for this assignment can be downloaded from the course web site:  

Dataset: Activity monitoring data [52K]  

The variables included in this dataset are:    
. steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)  
. date: The date on which the measurement was taken in YYYY-MM-DD format  
. interval: Identifier for the 5-minute interval in which measurement was taken  

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.


## Loading and preprocessing the data

```{r processfile, echo=TRUE}
        activityData <- read.csv(unzip(zipfile='activity.zip'))
```


## What is mean total number of steps taken per day?

```{r dailyMean, echo=TRUE}
        library(ggplot2)
        stepDailyMean <- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE)
        hist(stepDailyMean, col='green', xlab='Total Number of Steps Each Day')
        dailyMean <- mean(stepDailyMean, na.rm = TRUE)
        dailyMedian <- median(stepDailyMean, na.rm = TRUE)
```
  
The daily mean of steps taken is `r dailyMean` and the daily median is `r dailyMedian`.  


## What is the average daily activity pattern?

```{r dailyPattern, echo=TRUE}
        intervalAverages <- aggregate(x=list(steps=activityData$steps), 
                                      by=list(interval=activityData$interval),
                                      FUN=mean, na.rm=TRUE)
                
        plot(intervalAverages$interval,intervalAverages$steps, 
             type="l", 
             xlab="Interval", 
             ylab="Number of Steps",
             main="Average Number of Steps per Interval")

        maxAverage <- intervalAverages[which.max(intervalAverages$steps),]
```
  
The 5 minute interval with the maximum average of steps takenacross the dataset is interval `r maxAverage$interval` with `r maxAverage$steps` steps.   


## Imputing missing values

```{r missingValues, echo=TRUE}
        missingValues <- sum(is.na(activityData))
```
  
Total number of missing values in the dataset is `r missingValues`.  

```{r fillingData, echo=TRUE}
        newActivityData <- activityData 

        for (i in 1:nrow(newActivityData)) {
                if (is.na(newActivityData$steps[i])) {
                        newActivityData$steps[i] <- intervalAverages$steps[intervalAverages$interval==newActivityData$interval[i]]
                }
        }
```
  
New dataset created to with missing step values replaced with the value from the repective mean of 5-minute interval.  
The new data set is called newActivityData.

```{r newplot, echo=TRUE}
        newStepDailyMean <- tapply(newActivityData$steps, newActivityData$date, sum)
        hist(newStepDailyMean, col='blue', xlab='Total Number of Steps Each Day')
        newDailyMean <- mean(newStepDailyMean)
        newDailyMedian <- median(newStepDailyMean)
```
  
The daily mean for the new dataset without missing values is `r newDailyMean` and the daily median for the new dataset is `r newDailyMedian`.  
The new mean and median values for the new dataset are higher than the mean and mediam from the orginal dataset as missing values have been replaced with the mean values of the respective interval mean.

## Are there differences in activity patterns between weekdays and weekends?

```{r dayOfWeek, echo=TRUE}
        library(timeDate)
        
        newActivityData$DayOfWeek <- factor((isWeekday(newActivityData$date, wday=1:5)), 
                                            labels=c("weekend","weekday"))
```

### Plotting the results of weekday and weekend activities

```{r DoWPlots, echo=TRUE}
        library(lattice)
        
        newAverages <- aggregate(steps ~ interval + DayOfWeek, newActivityData, FUN=mean)
        
        xyplot(newAverages$steps ~ newAverages$interval|newAverages$DayOfWeek, 
               layout=c(1,2), 
               type="l",
               main="Average number of Steps per Interval seperated \nby Weekday and Weekend Activities",
               xlab="Interval", 
               ylab="Steps"
               )
```
